---
title: "p8105_hw3_ao2671"
author: "ao2671"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(p8105.datasets)
data("instacart")
```
# Problem 1

## Instacart description

Instacart dataset describes the orders in the Instacart system. It has the following variables: `r instacart %>% tbl_vars()`.
There is a total of `r instacart %>% distinct(order_id) %>% count()` orders by `r instacart %>% distinct(user_id) %>% count()` buyers from `r instacart %>% distinct(department) %>% count()`.
Orders can have multiple products in them. The dataset also contains the information on the aisles and departments for products purchased.
There are `r instacart %>% distinct(product_id) %>% count()` unique products purchased. The most popular products purchased 
is `r instacart %>% count(product_name) %>% arrange(desc(n)) %>% filter(row_number()==1) %>% select(product_name) `,
most popular aisle is `r instacart %>%  count(aisle) %>% arrange(desc(n)) %>% filter(row_number()==1) %>% select(aisle)`
and the most popular department is
`r instacart %>% count(department) %>% arrange(desc(n)) %>% filter(row_number()==1) %>% select(department)`. 
Orders are placed from `r instacart %>% pull(order_hour_of_day) %>% min()` 
to `r instacart %>% pull(order_hour_of_day) %>% max()` hours of day.

There are `r instacart %>% distinct(aisle) %>% count()` aisles. The aisles with the most ordered items are `r instacart %>% count(aisle) %>% arrange(desc(n)) %>% filter(row_number()<5) %>% select(aisle)`

## Plotting the number of orders per aisle in aisles with more than 10,000 orders

```{r plot}
instacart %>% count(aisle) %>% filter(n>10000) %>%
mutate(aisle = factor(aisle),
       aisle = fct_reorder(aisle, n)) %>%
ggplot(aes(x = aisle, y = n)) +
geom_point() +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
labs(title = "Number of orders per aisle" )

```


Arrange aisles sensibly, and organize your plot so others can read it.


## Top 3 products within "baking ingredients","dog food care" and "packaged vegetables fruits" aisles.

```{r aisles}
instacart %>% filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) %>%
group_by(aisle) %>%
count(product_name) %>%
mutate(rank=min_rank(desc(n))) %>%
filter(rank<4)%>%
arrange(aisle,rank)  
```

## Mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week
```{r aisles}
instacart %>% filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>%
group_by(product_name,order_dow) %>%
summarize(mean_hour = mean(order_hour_of_day))  %>%
pivot_wider(
  names_from = order_dow,
  values_from = mean_hour
)  %>%
rename( "Mon" = "0", "Tue" = "1","Wed" = "2","Thu" = "3","Fri" = "4","Sat" = "5","Sun" = "6")
```

# Problem 2

```{r accel_data, message=FALSE}
accel_data <- read_csv("./data/accel_data.csv") %>%  
   rename_at(vars(matches("^activity.")), ~ str_remove(., "^activity.")) %>% 
    mutate(weekday_vs_weekend = case_when(
                               day == "Saturday" | day == "Sunday" ~ "weekend",
                               TRUE                                ~ "weekday"
                                        )
    ) %>%
    relocate(week, day_id, day, weekday_v_weekend ) %>%
    mutate_at(5:1144, as.numeric)
```

## Description of the accelerometer dataset
Accelerometer dataset describes the activity of a man along a course of three weeks. There are `r accel_data %>% nrow()` rows, each of them represents one day with activity for each second of the day. We added additional column weekday_vs_weekend to indicate if a day is a weekday or weekend.


```{r trends}
accel_data %>% 
  mutate(
    sum_day = rowSums(select(accel_data, 5:1144))
  ) %>% 
  select(week, day_id, day, weekday_v_weekend, sum_day) 
```

In each week, there is one day where the activity is minimal (Monday on Week 1 and Saturdays on Week 2 and 3). It seems that for Saturdays on Week 2 and 3 the activity was not tracked (1 per each minute).




